Estimating the magnetization distribution within rectangular rock samples

### 2016-06-15 manuscript/manuscript.tex
compatibiliza o texto sobre a area do sensor com o resto do manuscrito

Over the last decades, scanning magnetic microscopy techniques have 
been increasingly used in paleomagnetism and rock magnetism. 
Different from standard paleomagnetic magnetometers, scanning 
magnetic microscopes produce high-resolution maps of the vertical 
component of the magnetic induction field on a plane located over 
the sample. 
These high-resolution magnetic maps can be used for estimating the 
magnetization distribution within a rock sample by inversion.
Previous studies have estimated the magnetization distribution 
within rock samples by inverting the magnetic data measured on a 
single plane above the sample. Here we present a new spatial domain
method for inverting the magnetic induction measured on four 
planes around the sample in order to retrieve its internal 
magnetization distribution. We have presumed that the internal 
magnetization distribution of the sample varies along one of
its axes. Our method approximates the sample geometry
by an interpretation model composed of a one-dimensional 
array of juxtaposed rectangular prisms with uniform magnetization. 
The Cartesian 
components of the magnetization vector within each rectangular prism 
are the parameters to be estimated by solving a linear inverse problem. 
Tests with synthetic data show the performance of our method in 
retrieving complex magnetization distributions even in 
the presence of magnetization heterogeneities. 
We have also applied our method to invert experimentally measured 
magnetic data produced by a highly-magnetized synthetic sample 
that was manufactured in the laboratory.
The results show that, even in the presence of apparent position
noise, our method was able to retrieve the magnetization distribution
consistent with the anhysteric remanance magnetization
induced in the sample.

@@

@@

The Cartesian coordinates as well as the magnetic induction components
in the observed data vector $\mathbf{d}$ (equation \ref{eq:obs-data-vec}) 
are referred to the MCS (Figure \ref{fig:sample}).
However, the measurements are taken in a different local coordinate 
system (LCS) for each observation plane (Figures
\ref{fig:sample-planes-cross}b-e) and must be subsequently converted to 
the MCS (Figure \ref{fig:sample-planes-cross}a).
This coordinate transformation is needed because the usual 
magnetic sensors can measure the vertical component of the 
magnetic induction produced by the sample on a plane located over it.
For this reason, the magnetic data on the four observation planes 
(Figure \ref{fig:sample-planes}) are
obtained by successively rotating the sample through $90^{\circ}$ 
intervals around its major axis ($x$-axis in Figure \ref{fig:sample}).
By repeating this rotating procedure and maintaining the same 
distance $h$ between the observation plane and the sample surface, its 
possible to obtain the magnetic induction on the four observation planes 
(Figures \ref{fig:sample-planes}).
The LCS on each plane has axes $x^{\prime}$, $y^{\prime}$ and $z^{\prime}$ 
(Figure \ref{fig:sample-planes-cross}b-e), where the
$x^{\prime}$-axis coincides with the $x$-axis of the MCS.

Note that this rotating procedure provides, on each observation plane, 
the component of the magnetic induction along the $z^{\prime}$-axis
of the respective LCS (Figure \ref{fig:sample-planes-cross}b-e).
It is easier, however, converting the measured $z^{\prime}$-component
data, as well as its Cartesian coordinates, from the LCS's to the MCS.
The geometrical relationships between these coordinate systems are
shown in Figure \ref{fig:sample-planes-cross}.
Table \ref{tab:coordinate-transformations} shows how to perform
this coordinate transformation. Let's consider, for example, the measurements 
obtained on the observation plane $\alpha = 1$ (Figures \ref{fig:sample-planes}b
and \ref{fig:sample-planes-cross}c). According to this table, 
the opposite of the measured $z^{\prime}$-component data correspond to the 
$y$-component of the magnetic induction that would be measured in the MCS;
the opposite of the Cartesian coordinates along the $z^{\prime}$-axis
correspond to the Cartesian coordinates along the $y$-axis in the MCS;
the Cartesian coordinates along the $y^{\prime}$-axis
correspond to the Cartesian coordinates along the $z$-axis in the MCS. 
The other lines of Table \ref{tab:coordinate-transformations} contain
the relationships used to convert the data obtained on the other planes.
Hereafter, it is implicit that all quantities with prime ($^{\prime}$) 
are referred to the LCS's while all the quantities without prime 
($^{\prime}$) are referred to the MCS.

@@

@@

By presuming that the rectangular rock sample can
be approximated by our previously described 
interpretation model, we define the linear inverse
problem of estimating its internal magnetization
distribution as an optimization problem.
This optimization problem consists in 
estimating a specific parameter vector 
$\mathbf{m} = \mathbf{m}^{\dagger}$ (equation \ref{eq:param-vec}) 
that minimizes the goal function
\begin{equation}
\Gamma(\mathbf{m}) = \left[ 
\mathbf{d} - \mathbf{B}(\mathbf{m})
\right]^{\top}
\left[ 
\mathbf{d} - \mathbf{B}(\mathbf{m})
\right] \: ,
\label{eq:goal-function}
\end{equation}
where $\mathbf{d}$ and $\mathbf{B}(\mathbf{m})$ are,
respectively, the observed and predicted data vectors
(equations \ref{eq:obs-data-vec} and \ref{eq:pred-data-vec-matrix}).
The minimization of the goal function $\Gamma(\mathbf{m})$ (equation 
\ref{eq:goal-function}) is equivalent to estimate the parameter vector 
minimizing the difference between the observed and predicted data 
vectors in the least-squares sense.

The parameter vector $\mathbf{m} = \mathbf{m}^{\dagger}$ minimizing 
the goal function (equation \ref{eq:goal-function}) is given by
\begin{equation}
\mathbf{m}^{\dagger} = \left( \mathbf{M}^{\top}\mathbf{M} \right)^{-1}
\mathbf{M}^{\top}\mathbf{d} \: ,
\label{eq:estimate}
\end{equation}
where $\mathbf{M}$ is the partitioned matrix defined in equation 
\ref{eq:sensitivity-matrix}.
We consider that the estimated parameter vector $\mathbf{m}^{\dagger}$
given by equation \ref{eq:estimate} approximates the real magnetization 
distribution within the rectangular rock sample. However, solutions given by the equation \ref{eq:estimate} can present non-uniqueness and instability. To overcome this ill- conditioning problem we need to impose a \textit{priori} information. As mentioned before, this procedure is generally called regularization. Here we imposed a constraint using ''Smoothness" regularization, generally called first-order Tikhonov regularization. Then equation \ref{eq:estimate} is now equal to 
\begin{equation}
\mathbf{m}^{\dagger} = \left( \mathbf{M}^{\top}\mathbf{M} + \mu_0 \,  \mathbf{R}^T \mathbf{R}\right)^{-1}
\mathbf{M}^{\top}\mathbf{d} \: ,
\label{eq:estimate_regul}
\end{equation}
where $\mu_0$ is a positive number called parameter regularization and $\mathbf{R}$ is a matrix with elements $-1$ and $1$. For example, if we have a sample formed by two prisms ($P=2$), the matrix $\mathbf{R}$ can be written as

\begin{equation}
\mathbf{R} =\begin{bmatrix}
 1& 0 & 0 & -1 & 0 & 0\\ 
0& 1 & 0 & 0 & -1 & 0\\
 0& 0 & 1 & 0 & 0 & -1\\
\end{bmatrix}_{3P-3 \times 3P}.
\label{eq:matrix_smoothness}
\end{equation}

Generally, the matrix $\mathbf{R}$ is composed by $3P = 3$ rows and $3P$ columns. Each element of the matrix can be described as

\begin{equation}
\left[ \mathbf{R} \right]_{ij} = 
\begin{cases}
1 \: , \quad i = j \\
-1 \: , \quad j = i+3
\end{cases} \: ,
\label{eq:elements_matrix}
\end{equation}
where the others elements of the matrix are zero.

@@

>>>>

Over the last decades, scanning magnetic microscopy techniques have 
been increasingly used in paleomagnetism and rock magnetism. 
Different from standard paleomagnetic magnetometers, scanning 
magnetic microscopes produce high-resolution maps of the vertical 
component of the magnetic induction field (flux density) 
on a plane located over the sample. 
These high-resolution magnetic maps can be used for estimating the 
magnetization distribution within a rock sample by inversion.
Previous studies have estimated the magnetization distribution 
within rock samples by inverting the magnetic data measured on a 
single plane above the sample. Here we present a new spatial domain
method for inverting the magnetic induction measured on four 
planes around the sample in order to retrieve its internal 
magnetization distribution. We have presumed that the internal 
magnetization distribution of the sample varies along one of
its axes. Our method approximates the sample geometry
by an interpretation model composed of a one-dimensional 
array of juxtaposed rectangular prisms with uniform magnetization. 
The Cartesian 
components of the magnetization vector within each rectangular prism 
are the parameters to be estimated by solving a linear inverse problem. 
Tests with synthetic data show the performance of our method in 
retrieving complex magnetization distributions even in 
the presence of magnetization heterogeneities. 
We have also applied our method to invert experimentally measured 
magnetic data produced by a highly-magnetized synthetic sample 
that was manufactured in the laboratory.
The results show that, even in the presence of apparent position
noise, our method was able to retrieve the magnetization distribution
consistent with the anhysteric remanance magnetization
induced in the sample.

@@

In practical situations, however, we are not able to measure
the magnetic induction produced by the sample at a point.
Due to the finite size of the magnetic sensors, what is
actually measured is the magnetic induction averaged on
the area (or volume) of the magnetic sensor. In the 
subsection \ref{subsec:Sensor active area}, we present how
we deal with this averaging effect.

@@

The Cartesian coordinates as well as the magnetic induction components
in the observed data vector $\mathbf{d}$ (equation \ref{eq:obs-data-vec}) 
are referred to the MCS (Figure \ref{fig:sample}).
However, the measurements are taken in a different local coordinate 
system (LCS) for each observation plane (Figures
\ref{fig:sample-planes-cross}b-e) and must be subsequently converted to 
the MCS (Figure \ref{fig:sample-planes-cross}a).
This coordinate transformation is needed because the usual 
magnetic sensors senses the vertical component of the 
magnetic induction produced by the sample on a plane located over it.
For this reason, the magnetic data on the four observation planes 
(Figure \ref{fig:sample-planes}) are
obtained by successively rotating the sample through $90^{\circ}$ 
intervals around its major axis ($x$-axis in Figure \ref{fig:sample}).
By repeating this rotating procedure and maintaining the same 
distance $h$ between the observation plane and the sample surface, its 
possible to obtain the magnetic induction on the four observation planes 
(Figures \ref{fig:sample-planes}).
The LCS on each plane has axes $x^{\prime}$, $y^{\prime}$ and $z^{\prime}$ 
(Figure \ref{fig:sample-planes-cross}b-e), where the
$x^{\prime}$-axis coincides with the $x$-axis of the MCS.

Note that this rotating procedure provides, on each observation plane, 
the component of the magnetic induction along the $z^{\prime}$-axis
of the respective LCS (Figure \ref{fig:sample-planes-cross}b-e).
For convenience, let us denote the measured $z^{\prime}$-component
data as follows
\begin{equation}
\mathbf{d}^{\prime} = \left[
\begin{array}{c}
\mathbf{d}^{0}_{z^{\prime}} \vspace{0.05in} \\
\mathbf{d}^{1}_{z^{\prime}} \vspace{0.05in} \\
\mathbf{d}^{2}_{z^{\prime}} \vspace{0.05in} \\
\mathbf{d}^{3}_{z^{\prime}}
\end{array}
\right]_{4N \times 1} \: ,
\label{eq:obs-data-vec-LCS}
\end{equation}
where $\mathbf{d}^{\alpha}_{z^{\prime}}$, 
$\alpha = 0, 1, 2, 3$, are $N \times 1$ vectors whose $i$-th 
element is the $z^{\prime}$-component of the magnetic 
induction which is measured at the $i$th observation
point, $i = 1, ..., N$, on the plane $\alpha$ 
(Figure \ref{fig:sample-planes-cross}).
It is easier, however, converting the measured data
$\mathbf{d}^{\prime}$ (equation \ref{eq:obs-data-vec-LCS}), 
referred to the LCS's, to the observed data vector $\mathbf{d}$
(equation \ref{eq:obs-data-vec}), which is referred to the MCS.

The geometrical relationships between the LCS's and MCS are
shown in Figure \ref{fig:sample-planes-cross}.
Table \ref{tab:coordinate-transformations} shows how to perform
this coordinate transformation. Let's consider, for example, the measurements 
obtained on the observation plane $\alpha = 1$ (Figures \ref{fig:sample-planes}b
and \ref{fig:sample-planes-cross}c). According to this table, 
the opposite of the measured $z^{\prime}$-component data correspond to the 
$y$-component of the magnetic induction that would be measured in the MCS;
the opposite of the Cartesian coordinates along the $z^{\prime}$-axis
correspond to the Cartesian coordinates along the $y$-axis in the MCS;
the Cartesian coordinates along the $y^{\prime}$-axis
correspond to the Cartesian coordinates along the $z$-axis in the MCS. 
The other lines of Table \ref{tab:coordinate-transformations} contain
the relationships used to convert the data obtained on the other planes.
Hereafter, it is implicit that all quantities with prime ($^{\prime}$) 
are referred to the LCS's while all the quantities without prime 
($^{\prime}$) are referred to the MCS.

@@

Notice that the forward problem formulated in this subsection
describes the magnetic induction produced by the sample at
the observation points. However, as pointed out before, 
we are not able to measure the magnetic induction at a point,
but an averaged magnetic induction. In the following subsection,
we present a strategy for incorporating this averaging effect
into the forward problem described here.

\subsection{Sensor active area}
\label{subsec:Sensor active area}

As pointed out at the final of the subsection 
\ref{subsec:Observed data vector}, 
magnetic sensors do not measure the magnetic induction 
field at a point. They measure instead the field averaged 
over the sensor active area (or volume), which is equivalent to a 
convolution of the magnetic induction field with 
the sensor answer function \citep{roth1989, egli2000, lima2013}.
The sensor-to-sample distance and the effective area (or volume) 
are important limiting factors determining the spatial resolution 
of magnetic sensors.

The magnetic sensors used in paleomagnetic applications have
different active area (or volume). 
\citet{uehara2007, uehara2008} presented a magneto-impedance (MI) 
sensor whose cylindrical active volume has $36$ $\mu$m in diameter and 
$5500$ $\mu$m in length, which limits its spatial resolution
if compared with the giant magnetoresistance 
(GMR) sensor presented by \citet{hankard2009} and the magnetic 
tunnel junction (MTJ) sensor presented by \citet{lima2014}, for example.
These two sensors are considered planar and have active areas
of, respectively, $9$~$\mu$m~$\times$~$36$~$\mu$m and 
$\approx 4$~$\mu$m~$\times$~$2$~$\mu$m.
Finally, \citet{fong2005} presented a low-transition-temperature 
superconducting quantum interference device (low T$_{c}$ SQUID) sensor
which is also considered planar and has an active area with
diameter of $80$~$\mu$m.
This is currently one of the most powerful magnetic sensors.

In the following, we present an analysis of the averaging effect produced
by the finite size of sensors and also a strategy
for numerically incorporating this effect into the forward 
problem described in the subsection 
\ref{subsec:Interpretation model and forward problem}. 
For convenience, we limit our study to square-planar 
sensors which are parallel to the sample surface and have an
active area $A$ (Figure \ref{fig:b-mag-flux-approx}a).
Let us first assume that the sensor detects,
at an observation point located on the observation plane $\alpha$, 
the $z^{\prime}$-component of the magnetic induction (flux density) 
(Figure \ref{fig:sample-planes-cross}) that is 
threaded through the active area $A$ 
(light grey square in Figure \ref{fig:b-mag-flux-approx}a)
of the sensor.
The measured $z^{\prime}$-component at the point 
$(x^{\prime}, y^{\prime}, z^{\prime})$ on the plane
$\alpha$ (Figure \ref{fig:sample-planes-cross}) can be described as a 
two-dimensional convolution given by
\begin{equation}
\tilde{d}^{\prime}(x^{\prime}, y^{\prime}, z^{\prime}) = 
\int \limits_{-\infty}^{\infty} \int \limits_{-\infty}^{\infty}
w(\eta, \zeta) \; d^{\prime}(x^{\prime} - \eta, 
                             y^{\prime} - \zeta,
                             z^{\prime}) \, d\eta d\zeta \: ,
\label{eq:b-mag-flux}
\end{equation}
where $w(\eta, \zeta)$ is the sensor answer function \citep{egli2000}. 
Here, we do not consider the
dependence with the sensor-to-sample distance $h$ 
(Figure \ref{fig:b-mag-flux-approx}a).
In this case, the sensor answer function 
is given by \citep{egli2000}
\begin{equation}
w(x^{\prime}, y^{\prime}) = \begin{cases}
\begin{split}
1/A &\: , \quad (x^{\prime}, y^{\prime}) \in A \\
0 &\: , \quad (x^{\prime}, y^{\prime}) \notin A
\end{split}
\end{cases} \: .
\label{eq:w-function}
\end{equation}
By substituting this sensor answer function into 
equation \ref{eq:b-mag-flux}, we obtain
\begin{equation}
\tilde{d}^{\prime}(x^{\prime}, y^{\prime}, z^{\prime}) = \frac{1}{A} 
\int \limits_{-\infty}^{\infty} \int \limits_{-\infty}^{\infty}
d^{\prime}(x^{\prime} - \eta, y^{\prime} - \zeta, z^{\prime}) \, d\eta d\zeta \: .
\label{eq:b-mag-flux2}
\end{equation}
Equation \ref{eq:b-mag-flux2} has an analytical solution
in the wavenumber domain \citep{egli2000}.
However, we opted for using an approximated solution in the
space domain. This approximation consists in discretizing the 
integral \ref{eq:b-mag-flux2} by using a midpoint rule on 
a regular grid of $Q_{x} \times Q_{y}$ points 
$(x^{\prime}_{j}, y^{\prime}_{j}) \in A$,
$j = 1, ..., Q$, where $Q = Q_{x} \, Q_{y}$
(open circles in Figure \ref{fig:b-mag-flux-approx}b).
According to this discretization, the active sensor area $A$ is
divided into a regular grid of square cells with area
$\Delta A$ (Figure \ref{fig:b-mag-flux-approx}b).
The approximated integral is given by
\begin{equation}
\tilde{d}^{\prime}_{i} \approx \frac{\Delta A}{A} \sum \limits_{j = 1}^{Q} d^{\prime}_{j} \: ,
\label{eq:b-mag-flux-approx}
\end{equation}
where $\tilde{d}^{\prime}_{i}$ is the measured 
$z^{\prime}$-component at the $i$th observation point
located at the center of the sensor active area $A$
(black circle in Figure \ref{fig:b-mag-flux-approx}b) and 
$d^{\prime}_{j}$ is the $z^{\prime}$-component at the
point $(x^{\prime}_{j}, y^{\prime}_{j})$, $j = 1, ..., Q$
(open circles in Figure \ref{fig:b-mag-flux-approx}b).
The $\tilde{d}^{\prime}_{i}$ measured on plane $\alpha$
is the $i$th element of the vector $\mathbf{d}^{0}_{z^{\prime}}$
(equation \ref{eq:obs-data-vec-LCS}).
Hence, by substituting $\Delta A = A/Q$ in equation {eq:b-mag-flux-approx}, 
we get
\begin{equation}
\tilde{d}^{\prime}_{i} \approx \frac{1}{Q} \sum \limits_{j = 1}^{Q} d^{\prime}_{j} \: .
\label{eq:b-mag-flux-approx2}
\end{equation}
Equation \ref{eq:b-mag-flux-approx2} shows that the measured
$z^{\prime}$-component $\tilde{d}^{\prime}_{i}$ 
at the $i$th observation point on a plane $\alpha$ is
approximately given by the magnetic induction
averaged over the active area $A$ of the sensor.

PAREI AQUI

@@

By presuming that the rectangular rock sample can
be approximated by our previously described 
interpretation model, we define the linear inverse
problem of estimating its internal magnetization
distribution as a constrained optimization problem.
This optimization problem consists in 
estimating a specific parameter vector 
$\mathbf{m} = \mathbf{m}^{\dagger}$ (equation \ref{eq:param-vec}) 
that minimizes the goal function
\begin{equation}
\Gamma(\mathbf{m}) = \left[ 
\mathbf{d} - \mathbf{B}(\mathbf{m})
\right]^{\top}
\left[ 
\mathbf{d} - \mathbf{B}(\mathbf{m})
\right] + 
\mu_0 \, \mathbf{m}^{\top} \mathbf{R}^{\top}\mathbf{R} \mathbf{m} \: ,
\label{eq:goal-function}
\end{equation}
where the superscript``$\top$" stands for transposition, 
$\mathbf{d}$ and $\mathbf{B}(\mathbf{m})$ are,
respectively, the observed and predicted data vectors
(equations \ref{eq:obs-data-vec} and \ref{eq:pred-data-vec-matrix}),
$\mu_0$ is a positive number called regularization parameter and
$\mathbf{R}$ is a sparse matrix with some elements equal to $-1$ and $1$.
For example, if the interpretation model is formed by 
two prisms ($P=2$), the matrix $\mathbf{R}$ is given by
\begin{equation}
\mathbf{R} =\begin{bmatrix}
 1& 0 & 0 & -1 & 0 & 0\\ 
0& 1 & 0 & 0 & -1 & 0\\
 0& 0 & 1 & 0 & 0 & -1\\
\end{bmatrix}_{3 \times 6}.
\label{eq:matrix_smoothness}
\end{equation} 
In a general case, $\mathbf{R}$ is a $3P-3 \times 3P$ matrix
whose $ij$th-element is given by
\begin{equation}
\left[ \mathbf{R} \right]_{ij} = 
\begin{cases}
\begin{split}
1 &\: , \quad i = j \\
-1 &\: , \quad j = i+3
\end{split}
\end{cases} \: .
\label{eq:elements_matrix}
\end{equation}
In equation \ref{eq:goal-function}, the regularization parameter $\mu_0$ 
controls the trade-off between 
the first and second terms at the right side. 
The first term represents the data fit
and the second term is the well-known first-order Tikhonov regularization 
function \citep{aster2005}, which imposes
``smoothness" to the estimated parameters.
The least-squares estimate of the parameter vector 
$\mathbf{m} = \mathbf{m}^{\dagger}$ minimizing 
the goal function (equation \ref{eq:goal-function}) is given by
\begin{equation}
\mathbf{m}^{\dagger} = \left( \mathbf{M}^{\top}\mathbf{M} + \mu_0 \,  
\mathbf{R}^T \mathbf{R}\right)^{-1}
\mathbf{M}^{\top}\mathbf{d} \: .
\label{eq:estimate}
\end{equation}
We consider that the estimated parameter vector $\mathbf{m}^{\dagger}$
given by equation \ref{eq:estimate} approximates the real magnetization 
distribution within the rectangular rock sample.

@@

 \begin{figure}
 \noindent \includegraphics[width=20pc]{Figs/Fig4b_HQ.pdf}
 \caption{Schematic representation of the integral \ref{eq:Phi-mag-flux2}.
 (a) Magnetic sensor with active area $A$ (light grey square) located at
 a distance $h$ of the sample (grey prism). (b) Detailed representation
 of the sensor active area $A$, which is centered at the observation point
 $(x^{\prime}, y^{\prime})$ (black circle). The square active area is discretized into 
 a regular grid of $Q$ square cells (limited by the dashed lines). Each
 cell is centered at a point $(x^{\prime}_{j}, y^{\prime}_{j})$, 
 $j = 1, ..., Q$, (open circles) and has an area $\Delta A$ equal to $A/Q$.}
 \label{fig:b-mag-flux-approx}
 \end{figure}
